---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# newdataset

<!-- badges: start -->
<!-- badges: end -->

The aim of the 'dataset' package is to make tidy datasets easier to release, exchange and reuse. It organizes and formats data frame 'R' objects into well-referenced, well-described, interoperable datasets into release and reuse ready form.

## Installation

You can install the development version of newdataset like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

“Tidy datasets provide a standardised way to link the structure of a dataset (its physical layout) with its semantics (its meaning). [tidyr](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)” While the tidyverse packages have made it easier to structure dataset into a tidy format, the semantics of the data are often remain at such a minimal level that it does not allow a meaningful reuse of the dataset. Often the only information we have about the meaning of the dataset is a column heading as a variable name, and a sequential `row.number` as an observation identifier. There is no way we can find out if a data subject is observed multiple times. Without a unit of measure information we may add together miles and kilometers, kilograms and tons, euros and dollars. 

The aim of the dataset package is to improve the semantic infrastructure of tidy datasets beyond the current capabilities of the tidyverse packages.

```{r example, echo=FALSE}
library(newdataset)
data("iris_dataset")
iris_dataset
```
To foster the findability, interoperability and reusability of a dataset, we attach standard metadata to the R object itself. Other metadata packages add this information to a separate CSV or JSON file, however, there is no guarantee that the next user will receive an intact R object and a separate metadata file. Therefore we attach this information to the native R object.

```{r bibentry}
print(get_bibentry(iris_dataset), "Bibtex")
```

How can we work with the famous `iris` dataset, add new observations, or make calculations if we do not happen to know that the numeric variables are all measured in centimeters? Do we know exactly how to tell a `setosa` observation from a `virginica`, if we do not know that the `Species` variable refers to species in the _Iris_ genus, and `setosa` stands for _Iris setosa_?


Can you divide the sepal length variable with sepal width? Technically you can, because they have the same numeric class. 

```{r}
iris_dataset$Sepal.Length[1:3] / iris_dataset$Sepal.Width[1:3]
```
But does it makes sense? With most statistical datasets, you only know if adding, subtracting, dividing and multiplying makes sense if you know the unit of measure. You may want to compare centimetres with centimetres, and avoid adding GDP values measured in dollars to GDP values measured in euros or yen.

So, we ensure that the unit of the variables remains attached in the attributes:


```{r variablelabels, eval=FALSE}
#needs to be corrected!
var_label(iris_dataset, unlist=TRUE)
```
```{r units}
var_unit(iris_dataset$Sepal.Length)
```
```{r defintion, eval=FALSE}
var_definition(iris_dataset$Species)
```
```{r namespace, eval=FALSE}
var_namespace(iris_dataset$Species)
```

Last, but not least, how can we know that this `iris_dataset` is that `iris` dataset? What happened with the dataset since creation? We record and attach to the object data provenance information.

```{r provenance, echo = FALSE}
cat(provenance(iris_dataset))
```


