---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# newdataset

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/antaldaniel/dataset/graph/badge.svg)](https://app.codecov.io/gh/antaldaniel/dataset)
[![lifecycle](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Project Status: WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/dataset)](https://cran.r-project.org/package=dataset)
[![CRAN_time_from_release](https://www.r-pkg.org/badges/ago/dataset)](https://cran.r-project.org/package=dataset)
[![Status at rOpenSci Software Peer Review](https://badges.ropensci.org/553_status.svg)](https://github.com/ropensci/software-review/issues/553)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.10574908.svg)](https://zenodo.org/record/6950435#.YukDAXZBzIU)
[![devel-version](https://img.shields.io/badge/devel%20version-0.3.3.0003-blue.svg)](https://github.com/antaldaniel/dataset)
[![dataobservatory](https://img.shields.io/badge/ecosystem-dataobservatory.eu-3EA135.svg)](https://dataobservatory.eu/)
<!-- badges: end -->

The aim of the _dataset_ package is to make tidy datasets easier to release, exchange and reuse. It organizes and formats data frame R objects into well-referenced, well-described, interoperable datasets into release and reuse ready form.

> The dataset package was archived on CRAN due to a documentation error.  This is a significantly reworked pre-release for rOpenSci review. Please review the package [concept](https://dataset.dataobservatory.eu/articles/new_requirements.html), the way the package handles variable-level metadata with the [defined](https://dataset.dataobservatory.eu/articles/defined.html) class, and bibliographical data as added to the attributes to the [dataset_df](https://dataset.dataobservatory.eu/reference/dataset_df.html) class in an extended [bibentry](https://dataset.dataobservatory.eu/articles/bibentry.html) object.

## Installation

You can install the development version of new_dataset like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

“Tidy datasets provide a standardised way to link the structure of a dataset (its physical layout) with its semantics (its meaning). [tidyr](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)” 

While the tidyverse packages have made it easier to structure datasets into a tidy format, the semantics of these datasets are not sufficiently rich for data exchange or publication, and often, they pose a difficulty for the same user if they take on work with the same dataset after a longer time.

The dataset package aims to create extensions of the modernised data.frame, i.e., tibble that contains rich semantic information for data exchange and reuse. It adds standardised information to the attributes (metadata) of the columns and the data frame as a whole, and it supports standard XML serialisations for statistical data exchange.

```{r example, echo=FALSE}
library(dataset)
data("iris_dataset")
iris_dataset
```

Our `dataset_df` class adds a `utils::bibentry` object to the attributes of the data.frame, and fills it up with the standard unassigned values of DataCite, an important open data publication standard. 

```{r print}
print(get_bibentry(iris_dataset), "Bibtex")
```

We created the `defined` class, an extension of `haven::labelled` from tidyverse, that goes beyond adding variable labels to the columns of a tidy dataset, and value labels to the categorical variables. We add two more crucial metadata: a unit of measure and a definition, with the ability to use statistical and web exchange standards to provide such information with linked open data. 

```{r defined}
gdp <- defined(
    c(3897, 7365), 
    label = "Gross Domestic Product", 
    unit = "million dollars", 
    definition = "http://data.europa.eu/83i/aa/GDP")

attributes(gdp)
```
A semantic application or a human user can look up the definition at  [http://data.europa.eu/83i/aa/GDP](https://data.europa.eu/83i/aa/GDP).

Some packages, for example, dataspice, help to add further information about the semantics of the dataset as a whole and some of its contents, but they detach such information from the data.frame that contains the data. Our aim was to retain all the critical information to publish, find, extend, join the dataset with other information in the R object itself.

```{r}
saveRDS(iris_dataset, file = tempfile())
```



